---
output:
  pdf_document: default
  html_document: default
---






#Importing the dataset and removing the rows with na BMI and Hypertension values
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)

health_raw <- read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/HMO_data.csv")
health=health_raw %>% filter(!(is.na(bmi)))
health=health %>% filter(!(is.na(hypertension)))
sapply(health,function(x) sum(is.na(x)))
sapply(health,function(x) sum(is.null(x)))

str(health)
#health
```

#Creating a New column 
```{r}
threshold=quantile(health$cost,probs=(.75))
health$Expensive <- ifelse(health$cost>=threshold, 1, 0)
glimpse(health)
```

#Creating Histograms for every numeric Variable
```{r}
library(ggplot2)
#Age
s=health%>%group_by(Expensive,age) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=s$age
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Age")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Age")+ylab("Average Cost")
rm(s,a,b,c)

#BMI
s=health%>%group_by(Expensive,bmi) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=s$bmi
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(position = "dodge",stat='identity',width = .35)+xlab("BMI")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(position="dodge",stat='identity',width = .35)+xlab("BMI")+ylab("Average Cost")
rm(s,a,b,c)

#Children
s=health%>%group_by(Expensive,children) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=s$children
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Children")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Children")+ylab("Average Cost")
rm(s,a,b,c)

#hist(health$smoker)
s=health%>%group_by(Expensive,smoker) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=s$smoker
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Smoker")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Smoker")+ylab("Average Cost")
rm(s,a,b,c)


#Location_type
s=health%>%group_by(Expensive,location_type) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=(s$location_type)
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Location Type")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Location Type")+ylab("Average Cost")
rm(s,a,b,c)

#hist(health$education_level)
s=health%>%group_by(Expensive,education_level) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=(s$education_level)
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Education Level")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Education Level")+ylab("Average Cost")
rm(s,a,b,c)

#hist(health$yearly_physical)
s=health%>%group_by(Expensive,yearly_physical) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=s$yearly_physical
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Yearly Physical")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Yearly Physical")+ylab("Average Cost")
rm(s,a,b,c)

#hist(health$exercise)
s=health%>%group_by(Expensive,exercise) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=s$exercise
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Exercise")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Exercise")+ylab("Average Cost")

rm(s,a,b,c)

#hist(health$married)
s=health%>%group_by(Expensive,married) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=s$married
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Married")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Married")+ylab("Average Cost")
rm(s,a,b,c)

#hist(health$hypertension)
s=health%>%group_by(Expensive,hypertension) %>%summarise(Freq = sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=(s$hypertension)
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Hypertension")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Hypertension")+ylab("Average Cost")
rm(s,a,b,c)

#hist(health$gender)
s=health%>%group_by(Expensive,gender)%>%summarise(Freq = sum(cost),avg=sum(cost),avg=mean(cost))
a=s$Expensive
b=s$Freq
c=(s$gender)
d=s$avg
ggplot(s, aes(y=b,x=c,fill=a))+geom_bar(stat='identity')+xlab("Gender")+ylab("Total Cost")
ggplot(s, aes(y=d,x=c,fill=a))+geom_bar(stat='identity')+xlab("Gender")+ylab("Average Cost")
rm(s,a,b,c)


library(usmap)
library(ggplot2)
us_states=map_data("state")
health$location=tolower(health$location)
data_merge=health %>% group_by(location) %>% summarize(avg_cost=mean(cost),total_sum=sum(cost))
merged=merge(data_merge,us_states,by.x='location',by.y='region')
merged=merged %>% arrange(order)
map=ggplot(merged)+geom_polygon(aes(x=long,y=lat,group=group,fill=avg_cost),color="black")
map+coord_map()

map=ggplot(merged)+geom_polygon(aes(x=long,y=lat,group=group,fill=total_sum),color="black")
map+coord_map()

```


#Linear Regression
```{r}
#First Iteration
lm<-lm(Expensive~.,data=health)
summary(lm)

res=resid(lm)
plot(fitted(lm),res)+abline(0,0)

qqnorm(health$Expensive)


#Second Iteration
#lm2=lm(Expensive~age+bmi+smoker+exercise+yearly_physical+hypertension+gender, data=health)
#summary(lm2)
```



```{r}
rm(health)
library(tidyverse)
library(ggplot2)
library(dplyr)

health_raw <- read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/HMO_data.csv")
health=health_raw %>% filter(!(is.na(bmi)))
health=health %>% filter(!(is.na(hypertension)))
sapply(health,function(x) sum(is.na(x)))
sapply(health,function(x) sum(is.null(x)))

threshold=quantile(health$cost,probs=(.75))
health$cost <- ifelse(health$cost>=threshold, 1, 0)
glimpse(health)
```




#Replacing values with numeric values for Models
```{r}

#smoker
health$smoker<-str_replace_all(health$smoker,"no","0")
health$smoker<-str_replace_all(health$smoker,"yes","1")
#location_type
health$location_type<-str_replace_all(health$location_type,"Country","0")
health$location_type<-str_replace_all(health$location_type,"Urban","1")
#education_level
health$education_level<-str_replace_all(health$education_level,"No College Degree","0")
health$education_level<-str_replace_all(health$education_level,"Bachelor","1")
health$education_level<-str_replace_all(health$education_level,"Master","2")
health$education_level<-str_replace_all(health$education_level,"PhD","3")
#yearly_physical
health$yearly_physical<-str_replace_all(health$yearly_physical,"No","0")
health$yearly_physical<-str_replace_all(health$yearly_physical,"Yes","1")
#exercise
health$exercise<-str_replace_all(health$exercise,"Not-Active","0")
health$exercise<-str_replace_all(health$exercise,"Active","1")
#married
health$married<-str_replace_all(health$married,"Not_Married","0")
health$married<-str_replace_all(health$married,"Married","1")
#gender
# Make sure to re-code female first
health$gender<-str_replace_all(health$gender,"female","1")
health$gender<-str_replace_all(health$gender,"male","0")

glimpse(health)
#view(health)
#summary(health)

```


#Removing NA's and Changing the Column type to numerics

```{r}
library(imputeTS)
#install.packages("zoo")
library(zoo)
#delete.na <- function(DF, n) {
#  DF[rowSums(is.na(DF)) <= n,]
#}

health$smoker <- as.numeric(health$smoker)
health$location_type<- as.numeric(health$location_type)
health$education_level <- as.numeric(health$education_level)
health$yearly_physical <- as.numeric(health$yearly_physical)
health$married <- as.numeric(health$married)
health$gender <- as.numeric(health$gender)
health$exercise <- as.numeric(health$exercise)
health$bmi=as.numeric(health$bmi)
health$cost=as.factor(health$cost)


glimpse(health)
#view(health)
#summary(health)
#nrow(health)

```





#Partioning the Dataset for Training and Testing
```{r}
library(kernlab)
library(caret)
health$cost=as.factor(health$cost)
trainList <-  createDataPartition(y=health$cost,p=.75,list=FALSE)
trainset <- health[trainList,]
testset <- health[-trainList,]
```


#SVM Model
```{r}

library(cvms)
library(tibble)
svm <- train(as.factor(cost) ~ ., data=trainset, method="svmRadial",preProc=c("center","scale"))
svm
svmpred=predict(svm,testset)
table(svmpred, testset$cost)
sum(diag(table(svmpred,testset$cost)))/sum(table(svmpred,testset$cost))
confusionMatrix(svmpred, testset$cost)

ctable=as.table(matrix(c(4040,654,136,738),nrow=2,byrow=TRUE))

fourfoldplot(ctable, color = c("cyan", "pink"),
             conf.level = 0, margin = 1, main = "Confusion Matrix")
```

#Recursive Partitioning and Regression Trees
```{r}
library(e1071)
library(caret)
library(rpart)
library(ggplot2)
library(rpart.plot)
rpartmodel<-train(cost~age+bmi+smoker+exercise+gender,data=trainset,method="rpart")
rpart.plot(rpartmodel$finalModel)
rpartPred<-predict(rpartmodel,testset)
confM2<-confusionMatrix(rpartPred,testset$cost)
confM2
```
